---
title: "Weight Lifting Exercise Manner Prediction"
author: "hoaihuongbk"
date: "8/24/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

## Goal

The goal of my report is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. After that, use the prediction model to predict 20 different test cases.

## Data Pre-Processing

The training data for this report are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

```{r}
training <- read.csv("./pml-training.csv", na.strings = c("NA", "#DIV/0!", ""))
testing <- read.csv("./pml-testing.csv", na.strings = c("NA", "#DIV/0!", ""))
str(training)
```

You can see there are many column with NA in training data, we should remove them first.Also, remove 7 columns from begining, because we don't use it in this report.

```{r}
training <- training[, colSums(is.na(training)) == 0]
testing <- testing[, colSums(is.na(testing)) == 0]
training <- training[, -c(1:7)]
testing <- testing[, -c(1:7)]
dim(training)
dim(testing)
```

For validation test, we will split training data into 2 part with 60% percentage for train and 40% for validation.

```{r message=FALSE}
library(caret)
set.seed(280890)
inTrain <- createDataPartition(training$classe, p=0.6, list=FALSE)
train <- training[inTrain,]
validation <- training[-inTrain,]
dim(train)
dim(validation)
```

## Prediction Models

We will use some algorithms:

* Classification Tree
* Random Forest
* Generalized boosted regression (GBR)

For each algorithm, we use `train` dataset for train data and `validation` dataset for validation test.

### Classification Tree

```{r message=FALSE, warning=FALSE, cache=TRUE}
control <- trainControl(method = "cv", number = 5)
model_ct <- train(classe ~ ., data=train, method="rpart", trControl=control)
print(model_ct, digits = 4)
predict_ct <- predict(model_ct, validation)
confusion_ct <- confusionMatrix(validation$classe, predict_ct)
accuracy_ct <- confusion_ct$overall[1]
accuracy_ct
```

### Random Forest

```{r message=FALSE, warning=FALSE, cache=TRUE}
model_rf <- train(classe ~ ., data = train, method = "rf", trControl = control)
print(model_rf, digits = 4)
predict_rf <- predict(model_rf, validation)
confusion_rf <- confusionMatrix(validation$classe, predict_rf)
accuracy_rf <- confusion_rf$overall[1]
accuracy_rf
```

### Generalized boosted regression (GBR)

```{r message=FALSE, warning=FALSE, cache=TRUE}
model_gbm <- train(classe ~ ., data=train, method="gbm", trControl = control, verbose = FALSE)
print(model_gbm, digits = 4)
predict_gbm <- predict(model_gbm, validation)
confusion_gbm <- confusionMatrix(validation$classe, predict_gbm)
accuracy_gbm <- confusion_gbm$overall[1]
accuracy_gbm
```

## Result

Classificatiom Trees (`r round(accuracy_ct * 100, digits=2)`% accuracy), Random Forest (`r round(accuracy_rf * 100, digits=2)`% accuracy), Generalized Boosted Regression (`r round(accuracy_gbm * 100, digits=2)`% accuracy)

**Random Forest** is the most accurate algorithm for the dataset among the applied three algorithm.

## Testing

```{r message=FALSE, warning=FALSE}
predict_t <- predict(model_rf, testing)
plot(predict_t)
predict_t
```
